SSA based global code motion of identical computations

### 2017-04-15 acmart-pacmpl-template.tex
algos

\subsection{Liveness using DJ Graph}
DJ-Graph \cite{Sreedhar1996} is a data structure that augments the dominator
tree with join-edges to keep track of data-flow in a program. We use DJ-Graph
and Merge Sets to compute liveness of variables as illustrated in
\cite{das2012}.  It is very efficient for computing liveness and does not
require any bitvectors to be maintained for each basic block. The underlying
simplicity of liveness computation is due to SSA form, where the the values only
flow (from def to use) either through dominator edges or the join edges (where
we insert a PHI). The DJ-graph contains both these edges which allows for
computation of merge sets for each basic block i.e., a set of all basic blocks
where the values can flow from a particular basic block.

@@

\begin{verbatim}
  Compute if variable a is liveout from basic block n
  Input: Node N, Variable a

@@

The original algorithm presented in \cite{das2012} has mismatched types in terms
of uses and nodes because each node can have many uses. Also, while iterating on
the dominator of each user of a variable we may reach to the beginning of a
function, in that case the inner while loop needs to terminate. These two cases
were missing from the algorithm and we came across them during implementation.

>>>>

\subsection{Liveness using DJ Graph}
DJ-Graph \cite{Sreedhar1996} is a data structure that augments the dominator
tree with join-edges to keep track of data-flow in a program. We use DJ-Graph
and Merge Sets to compute liveness of variables as illustrated in
\cite{das2012}.  It is very efficient for computing liveness and does not
require any bitvectors to be maintained for each basic block. The underlying
simplicity of liveness computation is due to SSA form, where the the values only
flow (from def to use) either through dominator edges or the join edges (where
we insert a PHI). The DJ-graph contains both these edges which allows for
computation of merge sets for each basic block i.e., a set of all basic blocks
where the values can flow from a particular basic block. We have implemented
merge-set computation from DJ-graph as illustrated in \cite{das2005}. A simplified
version of merge set that we implemented is presented here:

\begin{verbatim}
// Return true if the merge set of source node of
// a visited J-edge (incoming edge of lnode) is not
// the subset of the merge set of lnode.
bool still_inconsistent(Node lnode, JEdges JE, Visited V, MergeSet M)
   for (all incoming edges to lnode) do
     Let e = Incoming edge
     if (e is in JE && e in V) then
       Let snode = Source Node of e
       if (M(snode)!(Subset) M(lnode)) then
         return true
       end if
     end if
   end for
return false
\end{verbatim}

\begin{verbatim}
// Compute merge set top-down in breadth first order.
bool constructMergeSet_1(BFSList B, JEdges JE, DomLevel DL)
repeat = False
// List of visited edges
Visited V
for n in B do
  for e in (all incoming edges to n) do
    if (e is in JE && e not in V) then
      V(e) = true
      Let snode = Source Node of e
      Let tnode = Target Node of e
      Let tmp = snode
      Let lnode = NULL
      while (DL(tmp) ≥ DL(tnode)) do
        Merge(tmp) = Merge(tmp) U Merge(tnode) U {tnode}
        lnode = tmp
        tmp = dom-parent(tmp)
      end while
      repeat = still_inconsistent(lnode, JE, Merge)
    end if
  end for
end while
return repeat

// Construct Merge set of each node in control-flow-graph G of a function.
void  constructMergeSet(CFG G) {
  B = Breadth First Order of G
  JE = JEdges of G
  DL = List of Path length (from root) of each node in G.

  do // Call until a fixed point is reached.
    Repeat = constructMergeSet_1(B, JE, DL);
  while (Repeat);
}
\end{verbatim}

@@

\begin{verbatim}
bool isLiveOutUsingMergeSet()
  // Compute if variable a is liveout from basic block n
  Input: Node N, Variable a

@@

The original algorithm presented in \cite{das2012} has mismatched types in terms
of uses and nodes because each node can have many instructions and hence many
uses. Also, while iterating on the dominator of each user of a variable we may
reach to the beginning of a function, in that case the inner while loop needs to
terminate. These two cases were missing from the algorithm and we came across
them during implementation.
